### 2015-08-27 qdpm_paper.tex
added all images and a table to the paper with some of the gap info

    One of the simplest and most well studied models that exhibits topological order is the
    toric code. The toric code is also an exactly solvable model. Unfortunately it is
    a difficult model to physically realize. Another class of models that are well studied are
    the quantum dimer models (QDM).  These models can exhibit topological order and are also
    easier to experimentally realize, for example some short range resonating valence bond
    systems can be approximated with a QDM that provides a low energy effective description of
    the frustrated spins. Some correlation functions in QDMs are solvable using Pfaffian models
    but in general they must be studied numerically.

@@

        In the QDPM Ising degrees of freedom live on the links of the lattice.
        The operator $\hat{\sigma}^x_l$ acts on these links and has 
        eigenvalues $\sigma_l = \pm 1$, where $\sigma_l$ is the
        Ising variables of the $l^{\mathrm{th}}$ link on the lattice. The number operator for a give link is
        defined as $\hat{n}_l \equiv (1+\hat{\sigma}_x)/2$ and counts the number of dimers on link $l$.
        An exactl local gauge symetry exsists on the lattice due to the contraint that either one or
        four dimers touch any give vertex. In terms of $n_l$ we can define an operator that counts the
        number of dimers touching a given vertex $v$ as $n_v = \sum_{l\in v} n_l$.

        the local gauge transformation $G_v$ acting on a give vertex $v$ for the quantum dimer model
        can be expressed as
        %
        \begin{equation}
            \label{}
            G_v=e^{i \alpha (n_v - n_0)}
        \end{equation}
        %
        where $n_0$ is the constrained number of dimers touching a vertex, $n_0=1$ in the QDM model,
        The operator $G_v$ then acts in such a
        way that only the physical states, states satisfying the local constraint 
        are invariant. For the QDM $G_v =
        e^{i\alpha (0)}$ and therefore $G_v$ leaves a physical state unchanged for any
        $\alpha$ making the local gauge symetry $U(1)$. The local gauge transformation that leaves
        only configurations contaiing dimers and pentamers invarient is
        %
        \begin{equation}
            \label{eqn:gauge_trans}
            G_v = e^{i (\hat{n}_v -1) 2\pi/3}    
        \end{equation}
        %
        For the QDPM $\alpha$ is must be $2\pi/3$ due to the relaxed contraint of the QDM
        model that alows for pentamers. Relaxing the contraint of the QDM brings the local gauge
        symetry from $U(1)$ to $Z(3)$.

@@

        We work specificly at the RK point at wich the ground state wave function
        is an equal weighted superposition of fully packed dimer pentamer configurations
        %
        \begin{equation}
            \label{}
            \left| \Psi \right\rangle = \frac{1}{\sqrt{N}}\sum_{\{C\}} \left| C \right\rangle
            ,
        \end{equation}
        %
        where $\{C\}$ is the set of fully packed configurations. The Hilbert space can be defined
        using the gauge trasformation in Eq.~\ref{eqn:gauge_trans} as all $\left| C \right\rangle$
        satisfying $G \left| C \right\rangle = \left| C \right\rangle$.

@@

        The pentamer terms in the Hamiltonian can be constructed with the dynamics shown in
        Fig.~\ref{fig:pentamer_moves}.  
        \begin{figure}[htpb]
            \centering
            \includegraphics[width=0.8\linewidth]{pentamer_moves.pdf}
            \caption{a) Horizontal criation an anihilation, b) horizontal move, c) and d) are two
            differen possible diagonal moves in the same direction.}
            \label{fig:pentamer_moves}
        \end{figure}
        Each of the  moves shown in Fig.~\ref{fig:pentamer_moves} can be
        rotated to produce the symetry related counterpart (eg. vertical creation and anihilation). 
        As an example the creation and anhilation term in the Hamiltonian is
        \begin{multline}
        H_{\mathrm{pentamer\ creation}} = t_{\mathrm{create}}
            \sum_{\bstr}
            \left(
                \left|
                    \hprs
                \right\rangle
                \left\langle
                    \hspr
                \right|
                +
                \mathrm{h.c.}
            \right)
            \\
            - v_{\mathrm{create}}
            \sum_{\bstr}
                \left(
                    \left|
                        \hprs
                    \right\rangle
                    \left\langle
                        \hprs
                    \right|
                    +
                    \left|
                        \hspr
                    \right\rangle
                    \left\langle
                        \hspr
                    \right|
                \right)
        \end{multline}
        where the factors $t_{\mathrm{create}}$ and $v_{\mathrm{create}}$ are the coefficients of
        the penamer pair creation and anihilation respectivly. At the RK point
        $t/v=1$ for each all the petamer terms. 
        Writing the kinetic and potential energy terms as done above for all the dynamics shown in
        Fig.~\ref{fig:pentamer_moves} and all of their symetry related counterparts comprise the
        the pentamer terms of the Hamiltonian.

@@

@@

    \begin{figure}[htpb]
        \centering
        \includegraphics[width=0.8\linewidth]{spacial_dmr_cor.pdf}
        \caption{(Note: This figure was made using equal probabilities of plaquette filps and
        pentamer moves) Spacial correlation function of parralel dimers for the two different
        sublattices distiguesed by their color (red, green).}
        \label{fig:spacial_dmr_cor}
    \end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Estimating the Gap}
    \subsection{Imaginary time vison correlations}
    \begin{figure}[htpb]
        \centering
        \includegraphics[width=0.8\linewidth]{z2_origin_vison_time_cor.pdf}
        \caption{Imaginary time $z_2$ \textit{single} vison correlations at the origin in the QDM model (red) and QDPM
        model (green) on an $8\times$ lattice. For the QDPM the dimer/pentamer move fraction is $0.9$.}
        \label{fig:name}
    \end{figure}

    \subsection{Imaginary time dimer correlations}
    \begin{figure}[htpb]
        \centering
        \includegraphics[width=0.8\linewidth]{dimer_origin_time_cor.pdf}
        \caption{Imaginary time dimer correlations of a single dimer in the QDM model (red) and QDPM
        model (green) on an $8\times$ lattice. For the QDPM the dimer/pentamer move fraction is $0.9$. }
        \label{fig:}

>>>>

    One of the simplest and most well studied models that exhibits topological order is the
    toric code, an exactly solvable model. Unfortunately it is
    a difficult model to physically realize. Another class of models that are well studied are
    the quantum dimer models (QDM).  These models can exhibit topological order and are also
    easier to experimentally realize, for example some short range resonating valence bond
    systems can be approximated with a QDM that provides a low energy effective description of
    the frustrated spins. Some correlation functions in QDMs are solvable using Pfaffian models
    but in general they must be studied numerically.

@@

        In the QDPM Ising degrees of freedom live on the links of the lattice.
        The operator $\hat{\sigma}^x_l$ acts on these links and has 
        eigenvalues $\sigma_l = \pm 1$, where $\sigma_l$ is the
        Ising variables of the $l^{\mathrm{th}}$ link on the lattice. The number operator for a give link is
        defined as $\hat{n}_l \equiv (1+\hat{\sigma}_x)/2$ and counts the number of dimers on link $l$.
        An exactly local gauge symmetry exists on the lattice due to the constraint that either one or
        four dimers touch any give vertex. In terms of $n_l$ we can define an operator that counts the
        number of dimers touching a given vertex $v$ as $n_v = \sum_{l\in v} n_l$.

        the local gauge transformation $G_v$ acting on a give vertex $v$ for the quantum dimer model
        can be expressed as
        %
        \begin{equation}
            \label{}
            G_v=e^{i \alpha (n_v - n_0)}
        \end{equation}
        %
        where $n_0$ is the constrained number of dimers touching a vertex, $n_0=1$ in the QDM model,
        The operator $G_v$ then acts in such a
        way that only the physical states, states satisfying the local constraint 
        are invariant. For the QDM $G_v =
        e^{i\alpha (0)}$ and therefore $G_v$ leaves a physical state unchanged for any
        $\alpha$ making the local gauge symmetry $U(1)$. The local gauge transformation that leaves
        only configurations containing dimers and pentamers invariant is
        %
        \begin{equation}
            \label{eqn:gauge_trans}
            G_v = e^{i (\hat{n}_v -1) 2\pi/3}    
        \end{equation}
        %
        For the QDPM $\alpha$ is must be $2\pi/3$ due to the relaxed constraint of the QDM
        model that allows for pentamers. Relaxing the constraint of the QDM brings the local gauge
        symmetry from $U(1)$ to $Z(3)$.

@@

        We work specifically at the RK point at which the ground state wave function
        is an equal weighted superposition of fully packed dimer pentamer configurations
        %
        \begin{equation}
            \label{}
            \left| \Psi \right\rangle = \frac{1}{\sqrt{N}}\sum_{\{C\}} \left| C \right\rangle
            ,
        \end{equation}
        %
        where $\{C\}$ is the set of fully packed configurations. The Hilbert space can be defined
        using the gauge transformation in Eq.~\ref{eqn:gauge_trans} as all $\left| C \right\rangle$
        satisfying $G \left| C \right\rangle = \left| C \right\rangle$.

@@

        The pentamer terms in the Hamiltonian can be constructed with the dynamics shown in
        Fig.~\ref{fig:pentamer_moves}.  
        \begin{figure}[htpb]
            \centering
            \includegraphics[width=0.8\linewidth]{pentamer_moves.pdf}
            \caption{a) Horizontal creation an annihilation, b) horizontal move, c) and d) are two
            different possible diagonal moves in the same direction.}
            \label{fig:pentamer_moves}
        \end{figure}
        Each of the  moves shown in Fig.~\ref{fig:pentamer_moves} can be
        rotated to produce the symmetry related counterpart (eg. vertical creation and annihilation). 
        As an example the creation and annihilation term in the Hamiltonian is
        \begin{multline}
        H_{\mathrm{pentamer\ creation}} = t_{\mathrm{create}}
            \sum_{\bstr}
            \left(
                \left|
                    \hprs
                \right\rangle
                \left\langle
                    \hspr
                \right|
                +
                \mathrm{h.c.}
            \right)
            \\
            - v_{\mathrm{create}}
            \sum_{\bstr}
                \left(
                    \left|
                        \hprs
                    \right\rangle
                    \left\langle
                        \hprs
                    \right|
                    +
                    \left|
                        \hspr
                    \right\rangle
                    \left\langle
                        \hspr
                    \right|
                \right)
        \end{multline}
        where the factors $t_{\mathrm{create}}$ and $v_{\mathrm{create}}$ are the coefficients of
        the pentamer pair creation and annihilation respectively. At the RK point
        $t/v=1$ for each all the pentamer terms. 
        Writing the kinetic and potential energy terms as done above for all the dynamics shown in
        Fig.~\ref{fig:pentamer_moves} and all of their symmetry related counterparts comprise the
        the pentamer terms of the Hamiltonian.

@@

        In the QDPM there is a conserved winding operator in a given topological sector
        \begin{equation}
            W_{x,y} = (N_A - N_B)\mod{3} = n
            .
        \end{equation}
        For $W_y$ ($W_x$), $N_A$ is the number of dimers crossed on a vertical (horizontal) cut and
        $n=0$, $1$, or $2$ depending on the topological sector. The winding operator commutes with
        the Hamiltonian and therefore share the same eigenvalues. Since there are three winding numbers in
        each direction then the ground state is ninefold degenerate.

@@

    \begin{figure}[htpb]
        \centering
        \includegraphics[width=0.8\linewidth]{spacial_dmr_cor.pdf}
        \caption{(Note: This figure was made using equal probabilities of plaquette flips and
        pentamer moves) Spacial correlation function of parallel dimers for the two different
        sublattices distinguished by their color (red, green).}
        \label{fig:spacial_dmr_cor}
    \end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Estimating the Gap}

    \begin{table}[htpb]
        \centering
        \caption{$\Delta \tau = 150$}
        \label{tab:label}
        \begin{tabular}{c c c}
            \hline\hline
         L            & 8x8 Gap  & 16x16  \\ 
            \hline
         frac 1.0     &          & \\
            \hline
         Dimer origin &  0.03230 & 0.00806 \\
         $Z_2$ origin &  0.00787 & 0.00198  \\
         $Z_3$ origin &  NA      & NA \\
            \hline
         frac 0.9     &          & \\
            \hline
         Dimer origin &  0.00419 & 0.00429 \\
         $Z_2$ origin &  0.00216 & 0.00129  \\
         $Z_3$ origin &  0.00346 & 0.00359 \\
            \hline
         frac 0.5     &          & \\
            \hline
         Dimer origin &          & 0.02049 \\
         $Z_2$ origin &          & 0.00614  \\
         $Z_3$ origin &          & 0.01927  
        \end{tabular}
    \end{table}

    \subsection{Imaginary time vison correlations}
    \begin{figure}[htpb]
        \centering
        \includegraphics[width=0.8\linewidth]{z2_origin_vison_time_cor.pdf}
        \caption{Imaginary time $z_2$ \textit{single} vison correlations at the origin in the QDM model (red) and QDPM
        model (green) on an $8\times8$ lattice. For the QDPM the dimer/pentamer move fraction is $0.9$.}
        \label{fig:name}
    \end{figure}

    \subsection{Imaginary time dimer correlations}
    \begin{figure}[htpb]
        \centering
        \includegraphics[width=0.8\linewidth]{dimer_origin_time_cor.pdf}
        \caption{Imaginary time dimer correlations of a single dimer in the QDM model (red) and QDPM
        model (green) on an $8\times8$ lattice. For the QDPM the dimer/pentamer move fraction is $0.9$. }
        \label{fig:}
